Supporting Information

### 2018-03-27 SupportingInformation.Rmd
filled in ext risk section

source("mtime.R") #A function needed for caching
source("getcopula.R")
source("OurBiCopSelect.R")
source("FittingCopula_selective_loc.R")
source("normal_cop_AICw.R")
source("aphid_data_calling.R")
source("plankton_north_sea_data_calling.R")
source("NonParamStat.R")
source("copsurrog2d.R")
source("TestStats_multivar_dataset.R")
source("skewness_multivar_dataset.R")
source("ExtremeTailDep.R")
```

@@

## Extinction risk
```{r fn_popsim_ml_D, echo=F}

# A function to generate dispersal matrix D (a square matrix : numlocs by numlocs)
Dispersal_mat<-function(numlocs,d){
  D<-matrix(0,numlocs,numlocs)
  # A companion matrix (delta) that indicates how "off" a diagonal is:
  mat <- matrix(seq_len(numlocs ^ 2), ncol = numlocs)
  delta <- row(mat) - col(mat)
  #print(delta)

  diag(D)<-(1-d)
  D[delta ==1 | delta==-1] <- d/2

  D[numlocs-1,numlocs]<-d
  D[2,1]<-d

  stopifnot(all(colSums(D)==1)) # this is a check
  return(D)
}
#-----------------------------------------------------------------------------------
#The simulator of the multi-location model with a dispersal matrix.
#-------Args--------------------------------------------------
#p0       A length numlocs vector holding initial populations
#ns       An numsims by numlocs by numsteps array of the epsilons, 
#           where numsteps is the number of time steps you want
#
#--------Output------------------------------------------------------
#A numsims by numlocs by numsteps+1 array of populations
#----------------------------------------------------------------------
popsim_ml_D<-function(p0,ns,D,r,K){
  numsims<-dim(ns)[1]
  numlocs<-dim(ns)[2]
  numsteps<-dim(ns)[3]

  #initialization
  res<-array(NA,c(numsims,numlocs,numsteps+1))
  res[,,1]<-rep(p0,each=numsims)

  #step the populations forward
  for (tct in 1:numsteps){

    #stochastic ricker model : if r=0 then back to LC model

    #growth rates based on the noise
    lam_sto<-exp(ns[,,tct])  #numsims by numlocs matrix
    lam_ricker<-exp(r*(1-(res[,,tct]/K)))

    #growth prior to dispersal
    res[,,tct+1]<-lam_ricker*lam_sto*res[,,tct] #numsims by numlocs matrix

    #after dispersal
    tres<-t(res[,,tct+1]) # numlocs by numsims matrix after taking transpose
    temp<-D%*%tres  # a matrix multiplication (numlocs by numlocs) * (numlocs by numsims)
    res[,,tct+1]<-t(temp)

    #see if there is extinction
    h<-res[,,tct+1]
    h[h<1]<-0
    res[,,tct+1]<-h
  }

# function to calculate extinction risk
extrisk<-function(sims){
  totpop<-apply(FUN=sum,X=sims,MARGIN=c(1,3))
  return(apply(FUN=sum,X=(totpop==0),MARGIN=2)/dim(sims)[1])
}
```

```{r fn_plotter_ext_risk, echo=F}
# function to optionally plot extinction risk agsinst time and give you back extinction risk after specified numsteps 
plotter_ext_risk<-function(numsims,numsteps,numlocs,d,r,K,ploton){
  D<-Dispersal_mat(numlocs = numlocs,d=d)
  ns1<-retd(n=numsteps*numsims,d=numlocs,rl=1)# a righttail dep matrix(numpoints by numlocs,     
#                                                      numpoints=numsteps*numsims)

  ns1<-array(ns1,c(numsteps,numsims,numlocs))# convert to an array (numsteps by numsims by numlocs)
  ns1<-aperm(ns1,c(2,3,1)) # convert to an array (numsims by numlocs by numsteps)
  pops1<-popsim_ml_D(p0=rep(25,numlocs),ns=ns1,D=D,r=r,K=K)
  risk_right<-extrisk(pops1)

  ns2<-(-ns1)
  pops2<-popsim_ml_D(p0=rep(25,numlocs),ns=ns2,D=D,r=r,K=K)
  risk_left<-extrisk(pops2)

  if(ploton==T){
   par(mfrow=c(2,1))
    plot(0:(length(risk_right)-1),risk_right,type='l',
         xlab='Time',ylab='Risk',col='blue',panel.first = grid())
    mtext("blue : right tail-dep")
    plot(0:(length(risk_left)-1),risk_left,type='l',xlab='Time',ylab='Risk',col='red',panel.first = grid())
    mtext("red : left tail-dep")
  }

  return(list(risk_right_after_numsteps=risk_right[numsteps+1],
              risk_left_after_numsteps=risk_left[numsteps+1]))

}

```{r fn_varying_d, echo=F}
varying_d<-function(numsims,numsteps,numlocs,r,K){
  risk_right<-c()
  risk_left<-c()
  d_seq<-seq(from=0,to=1,by=0.01)
  for(d in d_seq){
    riskrl<- plotter_ext_risk(numsims=numsims,numsteps = numsteps,numlocs = numlocs,d=d,r=r,K=K,ploton=F)
    risk_r<-riskrl$risk_right_after_numsteps
    risk_l<-riskrl$risk_left_after_numsteps
    risk_right<-c(risk_right,risk_r)
    risk_left<-c(risk_left,risk_l)
  }
  par(mfrow=c(1,2))
  plot(d_seq,risk_left,xlab='d',ylab='Risk_left',ylim=c(0,max(risk_left)),type="b",col="red",panel.first = grid())
  plot(d_seq,risk_right,xlab='d',ylab='Risk_right',ylim=c(0,max(risk_right)),type="b",col="blue",panel.first = grid())
  #mtext("red : left tail-dep, blue : right tail-dep")
}
```

### LC model :
```{r fig_risk_r0_numlocs5, echo=F, fig.cap="LC model : Plot Extinction risk vs. d(degree of dispersal) for numlocs=5, numsteps=25, numsims=1000.\\label{fig_risk_r0_numlocs5}", fig.height=4, fig.width=8}
varying_d(numsims=1000,numsteps =25,numlocs=5,r=0,K=50)
```

```{r fig_risk_r0_numlocs25, echo=F, fig.cap="LC model : Plot Extinction risk vs. d(degree of dispersal) for numlocs=25, numsteps=25, numsims=1000.\\label{fig_risk_r0_numlocs25}", fig.height=4, fig.width=8}
varying_d(numsims=1000,numsteps =25,numlocs=25,r=0,K=50)
```

@@

@@

```{r fig_risk_r1_K100, echo=F, fig.cap="Plot Extinction risk vs. d(degree of dispersal) for r=1, K=100, numlocs=10, numsteps=25, numsims=1000.\\label{fig_risk_r1_K100}", fig.height=4, fig.width=8}
varying_d(numsims=1000,numsteps =25,numlocs=10,r=1,K=100)
```

```{r fig_risk_r1_K200, echo=F, fig.cap="Plot Extinction risk vs. d(degree of dispersal) for r=1, K=200, numlocs=10, numsteps=25, numsims=1000.\\label{fig_risk_r1_K200}", fig.height=4, fig.width=8}
varying_d(numsims=1000,numsteps =25,numlocs=10,r=1,K=200)
```

```{r fig_risk_r1_numlocs5, echo=F, fig.cap="Plot Extinction risk vs. d(degree of dispersal) for r=1, K=100, numlocs=5, numsteps=25, numsims=1000.\\label{fig_risk_r1_numlocs5}", fig.height=4, fig.width=8}
varying_d(numsims=1000,numsteps =25,numlocs=5,r=1,K=100)
```

```{r fig_risk_numlocs25, echo=F, fig.cap="Plot Extinction risk vs. d(degree of dispersal) for r=1, K=100, numlocs=25, numsteps=25, numsims=1000.\\label{fig_risk_r1_numlocs25}", fig.height=4, fig.width=8}
varying_d(numsims=1000,numsteps =25,numlocs=25,r=1,K=100)
```

  * With increasing $K$, extinction risk decreases [for a fixed $d$ , $r$ and $numlocs$] but this effect is more prominent in
case of left tail-dep data. [see Figure \ref{fig_risk_r1_K100} - Figure \ref{fig_risk_r1_K200}]

  * With increasing $numlocs$, extinction risk decreases [for a fixed $d$ , $r$ and $K$] but this effect is more prominent in
case of left tail-dep data. [see Figure \ref{fig_risk_r1_numlocs5} - Figure \ref{fig_risk_r1_numlocs25}]

>>>>

source("mtime.R") #A function needed for caching
source("getcopula.R")
source("OurBiCopSelect.R")
source("FittingCopula_selective_loc.R")
source("normal_cop_AICw.R")
source("aphid_data_calling.R")
source("plankton_north_sea_data_calling.R")
source("NonParamStat.R")
source("copsurrog2d.R")
source("TestStats_multivar_dataset.R")
source("skewness_multivar_dataset.R")
source("ExtremeTailDep.R")
source("mydispersal.R")
```

@@

## Extinction risk

### LC model :

@@

```{r fig_risk_r0_numlocs5, echo=F, cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="LC model : Plot Extinction risk vs. d(degree of dispersal).\\label{fig_risk_r0_numlocs5}", fig.height=4, fig.width=8}
set.seed(seed)
varying_d(numsims=10000,numsteps =25,numlocs=5,r=0,K=100)
```

```{r fig_risk_r0_numlocs25, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="LC model : Plot Extinction risk vs. d(degree of dispersal).\\label{fig_risk_r0_numlocs25}", fig.height=4, fig.width=8}
set.seed(seed)
varying_d(numsims=10000,numsteps =25,numlocs=25,r=0,K=100)
```

@@

 *  For ricker model when extinction risk calculated after $25$ years, was plotted against each $d$ (=degree of dispersion) we varied, we found that extinction risk was higher for data came from left tail dependent environmental fluctuations than right tail dependent one upto $r=0.7$. For $r>0.7$, extinction risk is greater for data came from right tail dependent environmental fluctuations.

  * With increasing $K$ from $100$ to $200$, extinction risk in both case gets decreased as expected. [see Figure \ref{fig_risk_r0.05} - Figure \ref{fig_risk_r0.05_K200}]

 * With increasing $numlocs$ from $5$ to $25$, extinction risk decreased [for a fixed $d$] but this effect was more prominent in case of data came from left tail-dep environmental noise. [see Figure \ref{fig_risk_r1.25} - Figure \ref{fig_risk_r1.25_numlocs25}].

```{r fig_risk_r0.05, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs. d(degree of dispersal).\\label{fig_risk_r0.05}", fig.height=4, fig.width=8}
set.seed(seed)
r<-0.05
numlocs<-5
K<-100
varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K)
```

```{r fig_risk_r0.05_K200, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs. d(degree of dispersal).\\label{fig_risk_r0.05_K200}", fig.height=4, fig.width=8}
set.seed(seed)
r=0.05
numlocs<-5
K<-200
varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K)
```

```{r fig_risk_r0.25, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs. d(degree of dispersal).\\label{fig_risk_r0.25}", fig.height=4, fig.width=8}
set.seed(seed)
r=0.25
numlocs<-5
K<-100
varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K)
```

```{r fig_risk_r0.5, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs. d(degree of dispersal).\\label{fig_risk_r0.5}", fig.height=4, fig.width=8}
set.seed(seed)
r=0.5
numlocs<-5
K<-100
varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K)
```

```{r fig_risk_r0.6, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs. d(degree of dispersal).\\label{fig_risk_r0.6}", fig.height=4, fig.width=8}
set.seed(seed)
r=0.6
numlocs<-5
K<-100
varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K)
```

```{r fig_risk_r0.7, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs. d(degree of dispersal).\\label{fig_risk_r0.7}", fig.height=4, fig.width=8}
set.seed(seed)
r=0.7
numlocs<-5
K<-100
varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K)
```

```{r fig_risk_r0.75, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs. d(degree of dispersal).\\label{fig_risk_r0.75}", fig.height=4, fig.width=8}
set.seed(seed)
r=0.75
numlocs<-5
K<-100
varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K)
```

```{r fig_risk_r0.9, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs. d(degree of dispersal).\\label{fig_risk_r0.9}", fig.height=4, fig.width=8}
set.seed(seed)
r=0.9
numlocs<-5
K<-100
varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K)
```

```{r fig_risk_r1.0, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs. d(degree of dispersal).\\label{fig_risk_r1.0}", fig.height=4, fig.width=8}
set.seed(seed)
r=1.0
numlocs<-5
K<-100
varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K)
```

```{r fig_risk_r1.25, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs. d(degree of dispersal).\\label{fig_risk_r1.25}", fig.height=4, fig.width=8}
set.seed(seed)
r=1.25
numlocs<-5
K<-100
varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K)
```

```{r fig_risk_r1.25_numlocs25, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs. d(degree of dispersal).\\label{fig_risk_r1.25_numlocs25}", fig.height=4, fig.width=8}
set.seed(seed)
r=1.25
numlocs<-25
K<-100
varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K)
```

```{r fig_risk_r1.5, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs. d(degree of dispersal).\\label{fig_risk_r1.5}", fig.height=4, fig.width=8}
set.seed(seed)
r=1.5
numlocs<-5
K<-100
varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K)
```

```{r fig_risk_r1.75, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs. d(degree of dispersal).\\label{fig_risk_r1.75}", fig.height=4, fig.width=8}
set.seed(seed)
r=1.75
numlocs<-5
K<-100
varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K)
```

```{r fig_risk_r1.95, echo=F,  cache=T, cache.extra=list(seed,mtime("mydispersal.R"),mtime("ExtremeTailDep.R")), fig.cap="Ricker model : Plot Extinction risk vs. d(degree of dispersal).\\label{fig_risk_r1.95}", fig.height=4, fig.width=8}
set.seed(seed)
r=1.95
numlocs<-5
K<-100
varying_d(numsims=10000,numsteps =25,numlocs=numlocs,r=r,K=K)
```
